---
title: "Reproducible research - Week 2 coursework"
output: github_document
---

```{r setup, include=FALSE, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)

```

## Loading and preprocessing the data

```{r,echo=TRUE}
temp <- tempfile()
download.file("http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip",temp)
unzip(temp)
unlink(temp)
activity <- read.csv("activity.csv")
activity_bkp<-activity
head(activity)


```

###### Remove incomplete cases

```{r, echo=TRUE}
activity_clean <- activity[complete.cases(activity),]
head(activity_clean)
```

###### Summary -- group steps by dates, library dplyr is used

```{r, echo=TRUE}
activity_summary_dates <- aggregate(steps~date,data=activity_clean,sum )

names(activity_summary_dates)<- c("dates","total_steps")
head(activity_summary_dates)
```

###### Histogram total steps vs day

```{r, echo=TRUE}
hist(activity_summary_dates$total_steps, xlab = "Total steps", col="green",ylab = "Days", main = "Total steps vs Days" )
```

###### Mean and median for group by dates data

```{r, echo=TRUE}
mean_steps <- mean(activity_summary_dates$total_steps)

median_steps <- median(activity_summary_dates$total_steps)

mean_steps

median_steps

```

##### Calculate mean and median for data grouped by interval and get the interval which has the max mean

```{r, echo=TRUE}
activity_summary_interval<-aggregate(steps~interval,data=activity_clean,mean)

names(activity_summary_interval)<- c("intervals","mean")

head(arrange(activity_summary_interval, desc(mean)),1)
```

##### Draw the plot, interval vs means of intervals 

```{r, echo=TRUE}
with(activity_summary_interval, {plot(intervals, mean, type="l",col=3,main ="Interval and means of intervals")})
ab <-abline(lm(activity_summary_interval$mean ~ activity_summary_interval$intervals))
```

##### Total rows with NA values 

```{r, echo=TRUE}
total_na<-sum(complete.cases(activity)==FALSE)
total_na
```

##### Replace incomplete cases
N/A replaces with mean value for steps

```{r, echo=TRUE}
activity$steps[is.na(activity$steps)] = mean(activity$steps, na.rm=TRUE)
head(activity,10)
```

##### Total steps, group by date, after replacing missing values

```{r, echo=TRUE}
activity_summary_dates_full <-  aggregate(steps~date,data=activity,sum )
names(activity_summary_dates_full)<-c("date","total_steps")
```

##### Histogram after completing missing values

```{r, echo=TRUE}
hist(activity_summary_dates_full$total, xlab = "Total steps", ylab = "Number of Days", main = "Total steps vs Number of days",col="blue")

hist(activity_summary_dates$total_steps, xlab = "Total steps", col="green",ylab = "Days", main = "Total steps vs Days", add=T)

legend("topright", c("Imputed", "Non-imputed"), col=c("blue", "green"), lwd=10)
```


###### Mean and median for group by dates data after imputation

```{r, echo=TRUE}
new_mean <- mean(activity_summary_dates_full$total_steps)

new_median <- median(activity_summary_dates_full$total_steps)

new_mean

new_median
```

##### Difference between the original mean, median and the values after impulation

```{r, echo=TRUE}
diff_mean<-new_mean -mean_steps

diff_mean

diff_median<-new_median -median_steps

diff_median

total_diff <- sum(activity$steps) - sum(activity_clean$steps)

total_diff

```

##### New Factor variable -Day

```{r, echo=TRUE}
weekdays <- c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")

activity$day = as.factor(ifelse(is.element(weekdays(as.Date(activity$date)),weekdays), "Weekday", "Weekend"))

table(activity$day)
```



##### Calculate mean data grouped by interval for imputed data

```{r, echo=TRUE}
steps_by_interval <- aggregate(steps ~ interval + day, activity, mean)

```


##### Plot mean vs intervals by Weekday and Weekend

```{r, echo=TRUE}
library(lattice)

xyplot(steps_by_interval$steps ~ steps_by_interval$interval|steps_by_interval$day, main="Average Steps per Day by Interval",xlab="Interval", ylab="Steps",layout=c(1,2), type="l")

```

